---
title: "Single-inoculation experiment: analyses of phenotypes"
author: "Rebecca Batstone"
date: "`r format(Sys.Date())`"
output: github_document
editor_options: 
  chunk_output_type: console
---

## Setup

### Global setup, load relevant packages

```{r setup, message=FALSE, warning = FALSE}
# Global setup
knitr::opts_chunk$set(echo = TRUE)

# Packages
library("tidyverse") ## includes ggplot2, dplyr, readr, stringr
library("car") ## Anova function
library("lme4") ## mixed effects models
library("knitr") ## produce the knitted doc
library("cowplot") ## produce paneled plots
library("DHARMa") ## residual diagnostics for glmm
library("RColorBrewer") ## color palettes
library("wesanderson") ## more color palettes
library("lmerTest") ## lmerModLmerTest
library("reshape2") ## dcast function
library("corrplot") ## to make correlation plots
library("psych") ## calculate p-vals of correlation matrix
```

### Load data

First, we need to read in and tidy the data

```{r Load data, warning=FALSE, message=FALSE}
plants_all <- read_csv("./Single-inoculation experiment/Data/single_evo_18Jan2018.csv", 
    col_types = cols(tray = col_character())) ## load data from single inoculation experiment

# make a unique plant ID
plants_all$plant_ID <- as.factor(do.call(paste, 
                                         c(plants_all[c("group","tray","plant")], sep = "-"))) 

# exclude controls for downstream analyses
plants <- droplevels(subset(plants_all, ! tray_type %in% c("control"))) 

# Load isolate data, with the original plant genotype they associated with
isolate_info <- read_csv("./Single-inoculation experiment/Data/isolate_info.csv")

# match isolate ID to origin, using isolate_info
plants$origin <- isolate_info$line_origin[match(plants$isolate, 
                                               isolate_info$isolate)] 

# make isolate ID a combination of origin and isolate 
plants$origin_isolate <- as.factor(paste0(plants$origin, "-", 
                                      plants$isolate))

# Determine whether isolates share an evolutionary history with plant line being assayed
plants$line_origin <- do.call(paste, c(plants[c("line","origin")], sep = "-")) # create a column combining "line" and "origin"
plants$history <- ifelse(grepl("anc1021", plants$line_origin, ignore.case = T), "anc1021",
         ifelse(grepl("anc1022", plants$line_origin, ignore.case = T), "anc1022",              
         ifelse(grepl("270-270", plants$line_origin, ignore.case = T), "Yes",
         ifelse(grepl("276-276", plants$line_origin, ignore.case = T), "Yes",
         ifelse(grepl("279-279", plants$line_origin, ignore.case = T), "Yes", 
         ifelse(grepl("267-267", plants$line_origin, ignore.case = T), "Yes",
         ifelse(grepl("313-313", plants$line_origin, ignore.case = T), "Yes",
                "No")))))))

# Make key variables factors
plants$history <- factor(plants$history, levels = c("anc1021","anc1022","No","Yes"))
plants$line <- as.factor(plants$line)
plants$group <- as.factor(plants$group)
plants$tray <- as.factor(plants$tray)
plants$isolate <- as.factor(plants$isolate)

plants_Em1022 <- droplevels(subset(plants, strain == "Em1022")) ## subset to Em1022 only

#str(plants_Em1022) ## 1150 plants
```

## Models

### How does the quality of rhizobia as symbionts differ between ancestral and derived strains? And does it depend on shared evolutionary history?

```{r Shoot_models, warning=FALSE, message=FALSE, dpi=600}
data.shoot <- subset(plants_Em1022, !is.na(shoot)) ## remove all trait values that are NAs

# Two models
# First model asks whether ancestral and derived isolates differ:
lm1 <- lmer(sqrt(shoot) ~ line + evo_state + (1|isolate) + (1|tray), data.shoot)
(lmm1_sum <- summary(lm1))
(ANODEV_shoot_state <- Anova(lm1, type=2)) ## no interaction, use type II SS

# Check model diagnostics
plot(lm1, resid(., scaled=TRUE) ~ fitted(.), abline = 0)
hist(residuals(lm1, type = "pearson" ), breaks = 100)
qqnorm(residuals(lm1))
ggplot(data.frame(lev=hatvalues(lm1),pearson=residuals(lm1,type="pearson")),
      aes(x=lev,y=pearson)) +
    geom_point() +
    theme_bw()
## all model diagnostics look ok

# test significance of random terms
lm1_isolate <- lmer(sqrt(shoot)~ line + evo_state + (1|tray), data.shoot)
lm1_tray <- lmer(sqrt(shoot)~ line + evo_state + (1|isolate), data.shoot)
(lm1_isolate_anova <- anova(lm1, lm1_isolate))
(lm1_tray_anova <- anova(lm1, lm1_tray))

# Subset the data to just the derived isolates
data.shoot.der = droplevels(subset(data.shoot, history != "anc1022")) 

# Second model: Fit a model with the fixed effect of line and isolate origin (home vs. away), 
## and isolate and group/tray (blocks in the greenhouse) as random effects
lm2 <- lmer(sqrt(shoot)~ line + history + (1|isolate) + (1|tray), data.shoot.der) 
## fit a line*comp interaction too, but it was not significant so it was removed

# Summarize model
(lmm2_sum <- summary(lm2))
(ANODEV_shoot_history <- Anova(lm2, type=2)) ## no interaction, use type II SS
## history significantly increases symbiont quality

# Check model diagnostics
plot(lm2, resid(., scaled=TRUE) ~ fitted(.), abline = 0)
hist(residuals(lm2, type = "pearson" ), breaks = 100)
qqnorm(residuals(lm2))
ggplot(data.frame(lev=hatvalues(lm2),pearson=residuals(lm2,type="pearson")),
      aes(x=lev,y=pearson)) +
    geom_point() +
    theme_bw()
## All model diagnostics look ok

# test significance of random terms
lm2_isolate <- lmer(sqrt(shoot)~ line + history + (1|tray), data.shoot.der)
lm2_tray <- lmer(sqrt(shoot)~ line + history + (1|isolate), data.shoot.der)
(lm2_isolate_anova <- anova(lm2, lm2_isolate))
(lm2_tray_anova <- anova(lm2, lm2_tray))
```

### How does the fitness of rhizobia differ between ancestral and derived strains? And does it depend on shared evolutionary history?

Next, we'll move to rhizobium fitness, measured by nodule number.

```{r Nod_models, warning=FALSE, message=FALSE, dpi=600}
data.nod <- subset(plants_Em1022, !is.na(nod) & nod > 0) ## remove all nod values that are NA or zero
hist(data.nod$nod) ## not normal, trying Poisson error distribution because it is count data
# Two models
## First model asks whether ancestral and derived isolates differ
glm1 <- glmer(nod ~ line + evo_state + (1|isolate) + (1|tray), data.nod, family="poisson",
              control=glmerControl(optimizer="bobyqa",
                           optCtrl=list(maxfun=1000000)))

# Model summary
(glmm1_sum <- summary(glm1))
(ANODEV_nod_state <- Anova(glm1, type=2)) ## no interaction, use type II SS

# Check residual diagnostics
simOut <- simulateResiduals(fittedModel = glm1, n = 1000)
plot(simOut)
deviance(glm1)/df.residual(glm1) ## close to 1, so good
testDispersion(simOut)

# test significance of random terms
glm1_isolate <- glmer(nod ~ line + evo_state + (1|tray), data.nod, family="poisson",
              control=glmerControl(optimizer="bobyqa",
                           optCtrl=list(maxfun=1000000)))
glm1_tray <- glmer(nod ~ line + evo_state + (1|isolate), data.nod, family="poisson",
              control=glmerControl(optimizer="bobyqa",
                           optCtrl=list(maxfun=1000000)))
(glm1_isolate_anova <- anova(glm1, glm1_isolate))
(glm1_tray_anova <- anova(glm1, glm1_tray))

# Subset the data to just the derived isolates
data.nod.der = droplevels(subset(data.nod, history != "anc1022"))
# Model 2: Fit GLMM with Poisson error distribution
glm2 <- glmer(nod ~ line*history + (1|isolate) + (1|tray), data.nod.der, family="poisson",
              control=glmerControl(optimizer="bobyqa",
                                   optCtrl=list(maxfun=1000000)))

# Model summary
(glmm2_sum <- summary(glm2)) ## there is a significant interaction for line 270 and evolutionary history, 
## so keeping the line*history interaction
(ANODEV_nod_history <- Anova(glm2, type=3)) ## interaction, use type III SS

## Check residual diagnostics
simOut <- simulateResiduals(fittedModel = glm2, n = 1000)
plot(simOut)
deviance(glm2)/df.residual(glm2) ## close to 1, so good
testDispersion(simOut)

# test significance of random terms
glm2_isolate <- glmer(nod ~ line*history + (1|tray), data.nod.der, family="poisson",
              control=glmerControl(optimizer="bobyqa",
                           optCtrl=list(maxfun=1000000)))
glm2_tray <- glmer(nod ~ line*history + (1|isolate), data.nod.der, family="poisson",
              control=glmerControl(optimizer="bobyqa",
                           optCtrl=list(maxfun=1000000)))
(glm2_isolate_anova <- anova(glm2, glm2_isolate))
(glm2_tray_anova <- anova(glm2, glm2_tray))
```

## Figures

### Figure 2 

Visualize how rhizobia quality and fitness depend on the match between rhizobia and host genotypes.

```{r fig2_SI_exp, warnings = FALSE}
# summarize for isolates
means_line_history <- plants_Em1022 %>% 
  group_by(line, history) %>% 
  summarize(n=n(), mean_shoot = mean(shoot, na.rm=TRUE), 
            mean_nod = mean(nod, na.rm=TRUE), 
            sd_shoot = sd(shoot, na.rm=TRUE), 
            sd_nod = sd(nod, na.rm=TRUE), 
            se_shoot = sd_shoot/sqrt(n), 
            se_nod = sd_nod/sqrt(n)) %>%
  as.data.frame(.)

levels(means_line_history$history) <- c("1022", "No", "Yes")
means_line_history$line <- factor(means_line_history$line, levels=c("270", "276", "279", "313", "267")) ## order plant lines
palette <- "Royal1"

#Make figures
(figa <- ggplot(data=subset(means_line_history, history != "1022"), aes(x=history, y=mean_shoot, color=history))+
  geom_line(aes(group=line), color="black", linetype="dotted", alpha=0.8)+
  geom_point(size=2)+
  geom_errorbar(aes(x=history, ymin=mean_shoot-se_shoot, ymax=mean_shoot+se_shoot), width=0.2, alpha=0.8)+
  #geom_text(aes(x=history, y=7, label=n))+
  facet_grid(~line)+
  xlab("Shared evolutionary history")+
  ylab("Rhizobium quality (shoot biomass, mg)")+
  guides(color=FALSE)+
  scale_color_manual(values=wes_palette(palette))+
  theme_bw() +  
  theme(
    panel.background = element_rect(fill=wes_palette(palette)[[3]]), 
    strip.background = element_rect(fill=wes_palette(palette)[[4]]), 
    axis.title = element_text(size=12, face = "bold"),
    axis.text = element_text(size=10),
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(),
    strip.text =  element_text(face = "bold", size = 12)
    )
)

(figb <- ggplot(data=subset(means_line_history, history != "1022"), aes(x=history, y=mean_nod, color=history))+
  geom_line(aes(group=line), color="black", linetype="dotted", alpha=0.8)+
  geom_point(size=2)+
  geom_errorbar(aes(x=history, ymin=mean_nod-se_nod, ymax=mean_nod+se_nod), width=0.2, alpha=0.8)+
  #geom_text(aes(x=history, y=3, label=n))+
  facet_grid(~line)+
  xlab("Shared evolutionary history")+
  ylab("Rhizobium fitness (nodule number)")+
  guides(color=FALSE)+
  scale_color_manual(values=wes_palette(palette))+
  theme_bw() +    
  theme(
    panel.background = element_rect(fill=wes_palette(palette)[[3]]), 
    strip.background = element_rect(fill=wes_palette(palette)[[4]]), 
    axis.title = element_text(size=12, face = "bold"),
    axis.text = element_text(size=10),
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(),
    strip.text =  element_text(face = "bold", size = 12)
    )
)
  
fig <- plot_grid(figa, figb, ncol=2, labels=c("A", "B")) # Combine figure panels
save_plot("./Single-inoculation experiment/Figures/Figure2.pdf", fig, dpi = 1000, base_width=7) #Save plot
```

### Calculate raw and standardized means for GWAS analyses

```{r means}
# means across all environments
raw_means_all <- plants_Em1022 %>% 
  group_by(isolate, origin) %>% 
  summarize(n=n(), mean_shoot = mean(shoot, na.rm=TRUE), 
            mean_nod = mean(nod, na.rm=TRUE), 
            sd_shoot = sd(shoot, na.rm=TRUE), 
            sd_nod = sd(nod, na.rm=TRUE), 
            se_shoot = sd_shoot/sqrt(n), 
            se_nod = sd_nod/sqrt(n)) %>%
  as.data.frame(.)

raw_means_all$line <- "All"

# get isolate ID, based on shoot rank
raw_means_all_rank <- raw_means_all %>% 
  filter(origin != "anc1022") %>%
  mutate(iso_ID = rank(mean_shoot))

# means within each environment
raw_means_line <- plants_Em1022 %>% 
  group_by(line, isolate, origin) %>% 
  summarize(n=n(), mean_shoot = mean(shoot, na.rm=TRUE), 
            mean_nod = mean(nod, na.rm=TRUE), 
            sd_shoot = sd(shoot, na.rm=TRUE), 
            sd_nod = sd(nod, na.rm=TRUE), 
            se_shoot = sd_shoot/sqrt(n), 
            se_nod = sd_nod/sqrt(n)) %>%
  as.data.frame(.)

# combine both:
raw_means <- rbind(raw_means_line, raw_means_all)

# add in iso_ID (based on shoot rank)
raw_means$iso_ID <- raw_means_all_rank$iso_ID[match(raw_means$isolate, 
                                               raw_means_all_rank$isolate)] 
# Add in shared evolutionary history
raw_means$line_origin <- do.call(paste, c(raw_means[c("line","origin")], sep = "-")) # create a column combining "line" and "origin"
raw_means$history <- ifelse(grepl("anc1021", raw_means$line_origin, ignore.case = T), "anc1021",
         ifelse(grepl("anc1022", raw_means$line_origin, ignore.case = T), "anc1022",              
         ifelse(grepl("270-270", raw_means$line_origin, ignore.case = T), "Yes",
         ifelse(grepl("276-276", raw_means$line_origin, ignore.case = T), "Yes",
         ifelse(grepl("279-279", raw_means$line_origin, ignore.case = T), "Yes", 
         ifelse(grepl("267-267", raw_means$line_origin, ignore.case = T), "Yes",
         ifelse(grepl("313-313", raw_means$line_origin, ignore.case = T), "Yes",
                "No")))))))

# Make key variables factors
plants$history <- factor(plants$history, levels = c("anc1021","anc1022","No","Yes"))

# Standardize means
stand_means <- raw_means %>%
  group_by(line) %>%
  mutate(stand_shoot = (mean_shoot - mean(mean_shoot))/sd(mean_shoot),
         stand_nod = (mean_nod - mean(mean_nod))/sd(mean_nod)) %>%
  as.data.frame(.)

# check means
stand_means %>% 
  group_by(line) %>%
  summarize(mean_shoot = mean(stand_shoot),
            sd_shoot = sd(stand_shoot),
            mean_nod = mean(stand_nod),
            sd_nod = sd(stand_nod)) ## mean of 0, sd of 1

# format for GEMMA (line as columns)
stand_means.w_shoot <- dcast(stand_means, isolate ~ line, value.var="stand_shoot")
colnames(stand_means.w_shoot)[2:7] <- paste("shoot", 
                                                colnames(stand_means.w_shoot[,c(2:7)]), sep = "_")
stand_means.w_nod <- dcast(stand_means, isolate ~ line, value.var="stand_nod")
colnames(stand_means.w_nod) <- paste("nod", colnames(stand_means.w_nod), sep = "_")
stand_means.w <- cbind(stand_means.w_shoot, stand_means.w_nod[,-1])                           
save(stand_means.w, file="./Single-inoculation experiment/Output/stand_means_for_GEMMA.Rdata")
```

### Figures S1 and S2

Heatmaps of isolate effects on each host

```{r FigS1_S2. fig.width = 10, fig.height = 16}
# heatmap variables:
stand_means$origin_iso <- do.call(paste, c(stand_means[c("origin","iso_ID")], sep = "-"))

# factor specifications
stand_means$line <- factor(stand_means$line, 
                                         levels = c("All", "270", "276",
                                                    "279", "313", "267"))
stand_means$origin <- factor(stand_means$origin, 
                                  levels = c("anc1022", "270",
                                              "276", "279", "313", "267"))

### order for figure
stand_means$origin_iso <- factor(stand_means$origin_iso,
                  levels = unique(stand_means$origin_iso[order(stand_means$origin,
                                                               stand_means$iso_ID)]))

stand_means.d <- filter(stand_means, origin_iso != "anc1022-NA")

### colour x-axis according to line
line_col <- c("black","#F8766D","#A3A500","#00BF7D","#00B0F6","#E76BF3")
names(line_col) <- levels(stand_means.d$line)

pdf("./Single-inoculation experiment/Figures/FigureS1.pdf", width = 10, height = 16)

(HM_shoot_IL<- ggplot(stand_means.d,
               aes(x = line, y = origin_iso, fill = stand_shoot))+
  geom_tile(color = "black") +
  xlab("Plant line isolate tested on") + 
  ylab("Isolate (plant line isolate evolved on-ID)") +
    
   scale_fill_gradient2(low = "blue",mid = "white", high = "red",
                      midpoint = 0,
                      space = "Lab",
                      name="Shoot biomass (g)") +
    
    # Label All column with isolate IDs (origin, overall rank)
    geom_text(data=subset(stand_means.d, line == "All"),
              aes(label = origin_iso), color = "black", size = 6) +
    theme_bw() +
    
    # add boxes around isolates that share an evolutionary history with the plant line being assayed
    annotate("rect", xmax = 2.5, ymax = 12.5, xmin= 1.5, ymin = 0.5, size = 2, colour="black", fill=NA)  + ## 270
    annotate("rect", xmax = 3.5, ymax = 15.5, xmin= 2.5, ymin = 12.5, size = 2, colour="black", fill=NA)  + ## 276
    annotate("rect", xmax = 4.5, ymax = 26.5, xmin= 3.5, ymin = 15.5, size = 2, colour="black", fill=NA)  + ## 279
    annotate("rect", xmax = 5.5, ymax = 32.5, xmin= 4.5, ymin = 26.5, size = 2, colour="black", fill=NA)  + ## 313
    annotate("rect", xmax = 6.5, ymax = 40.5, xmin= 5.5, ymin = 32.5, size = 2, colour="black", fill=NA)  + ## 267
    
    # Add asterisks to positive isolate effects when isolates share an evolutionary history with the plant line being assayed
    # geom_text(data=. %>%
    #           group_by(origin_iso) %>%
    #           filter(history == "Yes" & stand_shoot > 0),
    #           label = "*", size = 10, colour = "black", nudge_y = -0.3) +
 
    guides(colour=FALSE) +
    theme(axis.title.y = element_text(colour = "black", size = 20), 
        axis.text.y = element_blank(),
        axis.title.x = element_text(size=20), 
        axis.text.x = element_text(size=18, colour = line_col),
        axis.ticks = element_blank(),
        legend.title = element_blank(),
        legend.key.height = unit(1.2, "inch"),
        legend.text = element_text(size=16)))

dev.off()

HM_shoot_IL

pdf("./Single-inoculation experiment/Figures/FigureS2.pdf", width = 10, height = 16)

(HM_nod_IL<- ggplot(stand_means.d,
               aes(x = line, y = origin_iso, fill = stand_nod))+
  geom_tile(color = "black") +
  xlab("Plant line isolate tested on") + 
  ylab("Isolate (plant line isolate evolved on-ID)") +
  scale_fill_gradient2(low = "blue",mid = "white", high = "red",
                      midpoint = 0,
                      space = "Lab",
                      name="Nodules (no.)") +
    theme_bw() +
 
    # Label All column with isolate IDs (origin, overall rank)
    geom_text(data=subset(stand_means.d, line == "All"),
              aes(label = origin_iso), color = "black", size = 6) +
    
    # add boxes around isolates that share an evolutionary history with the plant line being assayed
    annotate("rect", xmax = 2.5, ymax = 12.5, xmin= 1.5, ymin = 0.5, size = 2, colour="black", fill=NA)  + ## 270
    annotate("rect", xmax = 3.5, ymax = 15.5, xmin= 2.5, ymin = 12.5, size = 2, colour="black", fill=NA)  + ## 276
    annotate("rect", xmax = 4.5, ymax = 26.5, xmin= 3.5, ymin = 15.5, size = 2, colour="black", fill=NA)  + ## 279
    annotate("rect", xmax = 5.5, ymax = 32.5, xmin= 4.5, ymin = 26.5, size = 2, colour="black", fill=NA)  + ## 313
    annotate("rect", xmax = 6.5, ymax = 40.5, xmin= 5.5, ymin = 32.5, size = 2, colour="black", fill=NA)  + ## 267
    
    # Add asterisks to positive isolate effects when isolates share an evolutionary history with the plant line being assayed
    # geom_text(data=. %>%
    #           group_by(origin_iso) %>%
    #           filter(history == "Yes" & stand_nod > 0),
    #           label = "*", size = 10, colour = "black", nudge_y = -0.3) +
 
    guides(colour=FALSE) +
    theme(axis.title.y = element_text(colour = "black", size = 20), 
        axis.text.y = element_blank(),
        axis.title.x = element_text(size=20), 
        axis.text.x = element_text(size=18, colour = line_col),
        axis.ticks = element_blank(),
        legend.title = element_blank(),
        legend.key.height = unit(1.2, "inch"),
        legend.text = element_text(size=16)))

dev.off()

HM_nod_IL
```

### Figure S3 

Correlation between rhizobia quality and fitness (across hosts)

```{r FigS3}
# Model for overall correlation:
# lm1
lm1 <- lm(mean_shoot ~ mean_nod, data=raw_means_all_rank)
summary(lm1)
Anova(lm1, type=2) # Significant positive correlation

# order factor levels
raw_means_all_rank$origin <- factor(raw_means_all_rank$origin, 
                                         levels = c("270", "276", "279", "313", "267"))

(plot_corr <- ggplot(raw_means_all_rank, aes(x=mean_nod, y=mean_shoot)) +
  geom_smooth(method=lm, se=TRUE, colour="black") +
  geom_errorbar(aes(ymin=mean_shoot-se_shoot, ymax=mean_shoot+se_shoot), colour="black") +
  geom_errorbarh(aes(xmin=mean_nod-se_nod, xmax=mean_nod+se_nod), colour="black") +
 
  # population aves
  geom_hline(data=raw_means_all %>%
               filter(origin == "anc1022"),
             aes(yintercept=mean(mean_shoot)), colour="black", linetype=2, size = 1) +
  geom_vline(data=raw_means_all %>%
               filter(origin == "anc1022"),
             aes(xintercept=mean(mean_nod)), colour="black", linetype=2, size = 1) +
  geom_hline(data=raw_means_all %>%
               filter(origin != "anc1022"),
             aes(yintercept=mean(mean_shoot)), colour="black", linetype=1, size = 1) +
  geom_vline(data=raw_means_all %>%
               filter(origin != "anc1022"),
             aes(xintercept=mean(mean_nod)), colour="black", linetype=1, size = 1) +
  
  geom_point(aes(colour=origin), size=4) +  
  annotate("text", x = 3, y = 16, label="italic(R) ^ 2 == 0.750", parse = TRUE, size=5) +  
  annotate("text", x = 3, y = 15.4, label='p < 0.001', size=5) +    
  theme_bw() + 
  xlab("Nodules (no.)") + 
  ylab("Shoot biomass (mg)") +
  geom_text(aes(label=iso_ID), size = 2) +  

  theme(axis.title.y = element_text(colour = "black", size = 20), 
        axis.text.y = element_text(size=18), 
        axis.title.x = element_text(colour = "black", size = 20), 
        axis.text.x = element_text(size=18), 
        legend.title = element_blank(),
        legend.text = element_text(colour="black", size=12, face="bold"),
        legend.box.background = element_blank(),
        legend.background = element_blank(),
        legend.position=c(0.9,0.22),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()))

save_plot("./Single-inoculation experiment/Figures/FigureS3.pdf", plot_corr, dpi = 1000,
          base_aspect_ratio = 1.3)
```

### Figure S4

Correlations across host environments. First, need to make wide format spreadsheets for each trait. Then, run corrplot on each.

```{r Figure S4}
# shoot biomass
stand_means.dw_shoot <- dcast(stand_means.d, iso_ID ~ line, value.var="stand_shoot")

# set line as row-names
stand_means.dw_shoot.m <- stand_means.dw_shoot[,-c(1:2)]
rownames(stand_means.dw_shoot.m) <- stand_means.dw_shoot[,1]

# set label colours
traitcolors1 <- c("#F8766D","#A3A500","#00BF7D","#00B0F6","#E76BF3")
names(traitcolors1) <- names(stand_means.dw_shoot.m)

# compute correlation matrix
stand_means.dw_shoot.mat <- cor(stand_means.dw_shoot.m, use="pairwise.complete.obs")

# compute the matrix of the p-value
source('./Single-inoculation experiment/Source code/corr_test.R')

p.mat_shoot <- cor.mtest(stand_means.dw_shoot.m) ## uncorrected
p.mat_shoot.c <- psych::corr.test(stand_means.dw_shoot.m, adjust="bonferroni")$p ## corrected

# Nodules

stand_means.dw_nod <- dcast(stand_means.d, iso_ID ~ line, value.var="stand_nod")

# set line as row-names
stand_means.dw_nod.m <- stand_means.dw_nod[,-c(1:2)]
rownames(stand_means.dw_nod.m) <- stand_means.dw_nod[,1]

# compute correlation matrix
stand_means.dw_nod.mat <- cor(stand_means.dw_nod.m, use="pairwise.complete.obs")

p.mat_nod <- cor.mtest(stand_means.dw_nod.m) ## uncorrected
p.mat_nod.c <- psych::corr.test(stand_means.dw_nod.m, adjust="bonferroni")$p ## corrected

## corrplot figure

# upper
bg_colors_upper1 <- p.mat_shoot < 0.05
bg_colors_upper1[bg_colors_upper1 == T] <- "grey"
bg_colors_upper1[bg_colors_upper1 == F] <- "white"
diag(bg_colors_upper1) <- "white" # if the diagonal values shall be white
# select colors from lower/upper1 part of matrix:
bg_colors_upper1 <- bg_colors_upper1[upper.tri(bg_colors_upper1, diag=T)]

# lower
bg_colors_lower1 <- p.mat_nod < 0.05
bg_colors_lower1[bg_colors_lower1 == T] <- "grey"
bg_colors_lower1[bg_colors_lower1 == F] <- "white"
diag(bg_colors_lower1) <- "white" # if the diagonal values shall be white
# select colors from lower/lower1 part of matrix:
bg_colors_lower1 <- bg_colors_lower1[lower.tri(bg_colors_lower1, diag=T)]

### and the plot
pdf("./Single-inoculation experiment/Figures/FigureS4.pdf", width = 10, height = 10, pointsize = 14)

par(xpd = TRUE)

corrplot(stand_means.dw_shoot.mat, method = "ellipse", type="upper",
         #order="hclust", 
         addCoef.col = "black",
         tl.pos = "tl",
         tl.col = traitcolors1,
         tl.srt=70,
         col = rev(brewer.pal(n = 8, name = "RdYlBu")),
         cl.lim=c(0,1),
         # p.mat = p.mat1.o, sig.level = 0.1, insig = "p-value",
         number.digits = 3,
         bg = bg_colors_upper1,
         diag = TRUE,
         mar=c(0,0,2,3))
         
corrplot(stand_means.dw_nod.mat, method = "ellipse", type="lower",
         #order="hclust", 
         addCoef.col = "black",
         col = rev(brewer.pal(n = 8, name = "RdYlBu")),
         add=T, 
         tl.pos="n", 
         cl.pos="n",
         # p.mat = p.mat1.o, sig.level = 0.1, insig = "p-value",
         number.digits = 3,
         bg = bg_colors_lower1,
         diag = TRUE)

dev.off()
```