---
title: "Variant filtering"
author: "Rebecca Batstone"
date: "`r format(Sys.Date())`"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, warning=FALSE, message = FALSE}
# packages
library("tidyverse") #includes ggplot2, dplyr, readr, stringr
```

## Filtering sites using vcftools

```{r vcftools_filter1, engine= 'bash', eval = FALSE}

$VCFTOOLS --vcf KH35c_rn.vcf --min-meanDP 20 --max-meanDP 230 --minQ 30 --max-missing 0.9 --max-non-ref-ac 48 --min-alleles 2 --max-alleles 2 --recode --recode-INFO-all --out KH35c_filt1
# kept 1330 out of a possible 21,247 Sites

# renamed
cp KH35c_filt1.recode.vcf ./filt1_1330.recode.vcf
```

Notes about filters: 

--min-meanDP 20
--max-meanDP 230 
Includes only sites with mean depth values (over all included individuals) greater than or equal to the "--min-meanDP" value and less than or equal to the "--max-meanDP" 

--minQ 30 
Includes only sites with Quality value above this threshold

--max-missing 0.9
Exclude sites on the basis of the proportion of missing data (defined to be between 0 and 1, where 0 allows sites that are completely missing and 1 indicates no missing data allowed). 
  * 0.9 therefore allows sites with 10% missing data

--max-non-ref-ac 48 
Include only sites with all Non-Reference (ALT) Allele Counts (ac) within the range specified, and including the specified value. 
  * Filters out any SNPs alternative to the reference, but homozygous across all my samples. 

--min-alleles 2 
--max-alleles 2
Include only sites with a number of alleles greater than or equal to the "--min-alleles" value and less than or equal to the "--max-alleles" value. 
  * I included only bi-allelic sites. (excluded 16 non-biallelic sites)

Notes on vcftools summary options:

--site-mean-depth
Generates a file containing the mean depth per site averaged across all individuals. This output file has the suffix ".ldepth.mean".

--hap-r2
Outputs a file reporting the r2, D, and D’ statistics using phased haplotypes. These are the traditional measures of LD often reported in the population genetics literature. The output file has the suffix ".hap.ld". This option assumes that the VCF input file has phased haplotypes.

--singletons
This option will generate a file detailing the location of singletons, and the individual they occur in. The file reports both true singletons, and private doubletons (i.e. SNPs where the minor allele only occurs in a single individual and that individual is homozygotic for that allele). The output file has the suffix ".singletons".

--site-quality
Generates a file containing the per-site SNP quality, as found in the QUAL column of the VCF file. This file has the suffix ".lqual".

## Pick representative singletons for each individual

I output singletons for KH35c_filt1.recode.vcf (487 singletons), and did the same for site quality. 

--exclude-positions
I determined how many singletons there were in the dataset (KH35c_filt1.vcf), singletons being a SNP that was only present in a single sample I used the vcftools output options described below. I chose a representative singleton for each individual by selecting the SNP within each individual that showed the highest quality score. Out of all the singletons (N=487), only 43 representatives were included in the filtered VCF (KH35c_filt2.vcf).

```{r remove_singletons}
# merge singletons and freq
singletons <- read_csv("./Data/1330_singletons.csv")
qual <- read_csv("./Data/1330_quality.csv")

SNP_info <- merge(x= singletons, y = qual, by = "POS")
# write.csv(SNP_info, "SNP_info.csv", row.names = FALSE)

# pick site with max quality for each individual
SNP_info_filt <- SNP_info %>%
  group_by(INDV) %>%
  filter(! QUAL == max(QUAL))
### 444 sites to exclude, 43 sites used as representatives for each sample

write.csv(SNP_info_filt, "./Output/SNP_info_exc.csv", row.names = FALSE)
```

### Use vcftools to filter out singletons

```{r filter_singletons, engine= 'bash', eval = FALSE}
# then, filtered out non-representative singletons (see note below, N = 437)
$VCFTOOLS --vcf KH35c_filt1.recode.vcf --exclude-positions SNPs_to_exclude.txt --recode --recode-INFO-all --out KH35c_filt2
# kept 886 out of a possible 1330 Sites
```

## Exclude SNPs in high (>0.95) LD

Checked pairwise LD for KH35c_filt2

Groups SNPs in high (>0.95) LD together (see high_LD_groups.xls for full details). Then picked the SNP from each group either showing the highest MAF or QUAL.

```{r LD_exclude}
LD_filt2 <- read_csv("./Data/886_LD.csv")

(LD_filt2_sum1 <- LD_filt2 %>%
  na.omit() %>%
  group_by(CHR) %>%
  summarize(mean_LD = mean(r2)))

LD_filt2$gen_dist <- abs(LD_filt2$POS2 - LD_filt2$POS1)

(LD_all_plot <- ggplot(LD_filt2, aes(x=gen_dist, y=r2, colour = CHR)) +
    geom_point() +
    theme_bw() +
    xlab("Genomic Span (Mbp)") + 
    ylab(expression("R"^"2")) +
    scale_color_discrete(name = "Genomic region",
      breaks = c("CP021825.1","CP021826.1","CP021827.1","CP021828.1"),
      labels=c("Chromosome","pAccA","pSymA","pSymB")) +
    #scale_x_continuous(limits = c(0,60)) +
    theme(axis.title.y = element_text(colour = "black", size = 18), 
        axis.text.y = element_text(size=16), 
        axis.title.x = element_text(size=18), 
        axis.text.x = element_text(size=16), 
        legend.position=c(0.8,0.8),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()))

# how many sites are at LD > 95%?
(LD_filt2_sum2 <- LD_filt2 %>%
  na.omit() %>%
  filter(r2 > 0.95) %>%
  summarize(count_LD = length(r2)))

prop_LD <- LD_filt2_sum2/length(LD_filt2$r2)

# what's the max distance between SNPs at r2 > 0.95
(LD_filt2_sum3 <- LD_filt2 %>%
  na.omit() %>%
  filter(r2 > 0.95) %>%
  summarize(max_dist = max(gen_dist)))

# LD groups: filter for SNPs @ LD > 0.95, group by site
LD_filt2_sum4 <- LD_filt2 %>%
  na.omit() %>%
  filter(r2 > 0.95) %>%
  group_by(POS1) %>%
  arrange(desc(POS2))
  
write.csv(LD_filt2_sum4, "./Output/high_LD_groups.csv", row.names = FALSE)

# plot of no vars (x), group count on y

LD_groups <- read_csv("./Data/high_LD_groups_plot.csv")

LD_group_sum <- LD_groups %>%
  group_by(LD_group) %>%
  summarize(mean_no_vars = mean(no_var_group))

LD_group_sum$mean_no_vars <- as.factor(LD_group_sum$mean_no_vars)

(LD_groups_plot <- ggplot(LD_group_sum, aes(x=mean_no_vars)) +
    geom_bar(stat="count") +
    theme_bw() +
    xlab("Variants per LD group (no.)") + 
    ylab("LD groups (no.)") +
    theme(axis.title.y = element_text(colour = "black", size = 18), 
        axis.text.y = element_text(size=16), 
        axis.title.x = element_text(size=18), 
        axis.text.x = element_text(size=16), 
        legend.position="none",
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()))

# merge spreadsheets
MAF_filt2 <- read_csv("./Data/MAF_LD_groups.csv")
qual_filt2 <- read_csv("./Data/QUAL_LD_groups.csv")
LD_groups_filt2 <- read_csv("./Data/high_LD_groups_UP.csv") # 135 SNPs

SNP_info2 <- merge(x= MAF_filt2, y = qual_filt2, by = "POS")
SNP_info3 <- merge(x= SNP_info2, y = LD_groups_filt2, by = "POS")

# filter variants within ea group, highest MAF and QUAL
SNP_info_filt2 <- SNP_info3 %>%
  group_by(group_no) %>%
  filter(MAF == max(MAF)) %>%
  filter(QUAL == max(QUAL))

exclude_pos <- SNP_info_filt2$POS # 31 SNPs, one representative from each group
SNP_info3_sub <- subset(SNP_info3, ! POS %in% exclude_pos) # 103 SNPs to exlude (see note below)
write.csv(SNP_info3_sub, "./Output/SNPs_to_exclude_LD.csv", 
          row.names = FALSE)
```

Note about LD groups: some SNPs appear in multiple groups. 288074, 288091 (both in groups 19 and 24), 365824 (in groups 11 and 19).

### Use vcftools to filter out linked variants

```{r filter_LD_SNPs, engine= 'bash', eval = FALSE}

$VCFTOOLS --vcf KH35c_filt2.recode.vcf --exclude-positions SNPs_to_exclude_LD.txt --maf 0.001 --recode --recode-INFO-all --out KH35c_final
# 363 out of a possible 886 Sites
cp KH35c_final.recode.vcf ../../GEMMA/GEMMA_363.recode.vcf
```

--maf 0.001
Include only sites with a Minor Allele Frequency greater than or equal to the "--maf" value
* Excludes homozygous sites with missing genotypes

